\chapter*{Foreword}

\section*{Description}

This updated paper presents an introduction to color management for motion-picture, television and games production, including visual effects, animation. It expands on the original’s focus on motion-picture visual effects and animation as a reflection of the expanded scope of digital imaging in the film and television production process, from capture to post-production to grading and final display, and as a reflection of the consistency in basic approaches used by visual effects, animation and increasingly games projects.

Color impacts many areas of the movie and the game-making process. From on-set capture to the many phases of the computer graphics pipeline: texture painting, look development, lighting, rendering, and compositing, to grading and finishing for the theater and home. Handling color is a tricky problem. This paper presents an introduction to color science, measuring and encoding color and core concepts like gamuts, transfer functions, and scene-referred and output-referred colorimetry. The paper then extends these concepts to their use in modern motion-picture and games production, including an introduction to efforts on digital color standardization in the motion-picture industry, ACES and ASC CDL, and how readers can experiment with all of these concepts for free using open-source software like OpenColorIO and Colour Science for Python.

\section*{Authorship}

The updated paper was authored by Alex Forsythe, Alex Fry, Stefan Luka, Thomas Mansencal, Kevin Shaw and Nick Shaw. 

The paper continues to include much of Jeremy Selan’s original text.

The update was edited by Haarm-Pieter Duiker.

The update was reviewed by members of the VES Technology Committee and others, including ...

The original paper was authored by Jeremy Selan and reviewed by the members of the VES Technology Committee including Rob Bredow, Dan Candela, Nick Cannon, Paul Debevec, Ray Feeney, Andy Hendrickson, Gautham Krishnamurti, Sam Richards, Jordan Soles, and Sebastian Sylwan.

\section*{On the web}

The VES Technology Committee has assembled an expert panel that can be reached at ves-tech-color@googlegroups.com for questions, comments, and errata related to this document.

Visit CinematicColor.org for additional resources related to motion-picture color management.
