\section{Software}
\label{sec:software}

This is a list of software applications mentioned elsewhere in the document. It is not intended to be comprehensive. Nor should it be taken as an endorsement of one particular application over another.

\subsection{Open Source Software}
\label{subsec:open-source-software}

DCRAW, Dave Coffin
Open source raw development software, giving greater control of image processing than is available in manufacturers’ software. It also permits bypassing of many of the ‘sweeteners’ often included, and thus allows decoding to a more pure scene-referred image. https://www.cybercom.net/~dcoffin/dcraw/

OpenColorIO
See Section 4.2 above
http://opencolorio.org

OpenImageIO
An open source library for reading, writing and manipulating image files, with particular emphasis on formats used in VFX, and not supported in standard image reader/writers. https://openimageio.org

\subsection{Commercial Software}
\label{subsec:commercial-software}
After Effects, Adobe

Baselight, FilmLight

CryEngine, Crytek

Flame, Autodesk

Fusion, Blackmagic Design

Lattice, Video Village
LUT creation and conversion application for Mac OS. https://lattice.videovillage.co

LightSpace, Light Illusion
Display calibration software for Windows, which is also able to create and manipulate LUTs. https://www.lightillusion.com

LiveGrade, Pomfort
Mac OS software for controlling LUT boxes to grade a live signal and save the grades and other metadata to be applied elsewhere. Works together with other software from Pomfort for managing metadata and creating dailies. https://pomfort.com/livegradepro

Lumberyard, Amazon

Lustre, Autodesk

Mistika, SGO

Nuke, Foundry

Nucoda, Digital Vision

On-Set Dailies, Colorfront

Prelight, FilmLight
Look creation software for Mac OS based around BLG (Baselight Saved Grade) but also capable of working in a more limited ASC CDL mode. Available in a free version which allows a DIT or Cinematographer to work with the full Baselight toolset to create looks using still images, and a paid version which can control LUT boxes for live grading and metadata management. https://www.filmlight.ltd.uk/products/prelight/overview_pl.php

Resolve, Blackmagic Design

Unity, Unity Technologies

Unreal Engine, Epic Games

